---
title: "R Notebook"
output: html_notebook
---

```{r}

library(tidyverse)
library(CodeClanData)

```

```{r}

students

```

```{r}
all_deaths
```

```{r}
nrow(all_deaths)
```
```{r}
ncol(all_deaths)
```
```{r}
dim(all_deaths) # rows and columns
```

```{r}
names(all_deaths)
```

```{r}
head(all_deaths)
```

```{r}
tail(all_deaths, 1)
```

```{r}
glimpse(all_deaths)
```

```{r}
summary(all_deaths)
```

```{r}
view(all_deaths)
```


### Select

```{r}
select(all_deaths, name, allegiances, gender, nobility, year_of_death) # pick which columns to pull from the data create new dataframe
```

```{r}
deaths_no_name <- select(all_deaths, -name) # use minus (or !) to discount columns
```

```{r}
select(all_deaths, contains("death"))
```

```{r}
select(all_deaths, year_of_death, book_of_death, death_chapter)
```

```{r}
select(all_deaths, book_of_death, book_intro_chapter, book1_GoT, book2_CoK, book3_SoS, book4_FfC, book5_DwD)
```

```{r}
select(all_deaths, contains("book"))
```

```{r}
select(all_deaths, 1,2,3) # picks columns by their number, is ORDER SPECIFIC 
```

```{r}
select(all_deaths, 1:5) # ranges also work
```

```{r}
select(all_deaths, name:book_of_death) # ranges also work with names of solumns if they are all in a row
```

```{r}
select(all_deaths, -c(name:book_of_death)) # also works with minus to exclude certain columnsranges also work with names of columns if they are all in a row
```


```{r}
select(all_deaths, contains("death"), contains("book"))
```




### filter

---------------------------------------------------------------------------------------------

```{r}

filter(all_deaths, allegiances == "Lannister") #filter where row equals a string

```

```{r}

filter(all_deaths, allegiances != "Lannister") #filter where row equals a string

```

```{r}
filter(all_deaths, allegiances %in% c("Lannister", "House Lannister")) # can use the %in% operator
```

```{r}
filter(all_deaths, year_of_death >= 299)
```

```{r}
filter(all_deaths, allegiances == "None")
```

```{r}
filter(all_deaths, allegiances == "Lannister" & nobility == 1) # multiple filter example with & sign
```

```{r}

filter(all_deaths, allegiances %in% c("Lannister", "House Stark"))

```

```{r}
filter(all_deaths, book_intro_chapter >= 5 & book_intro_chapter <= 10)
```

```{r}

filter(all_deaths, between(book_intro_chapter, 5, 10)) # same as above code uses function "between"

```

```{r}
filter(all_deaths, str_detect(allegiances, "Lannister"))
```

Find where the year_of_deathis less than or equal to 299.
Find the females (gender is 0) who are not Lannisters
Find just the data for the characters “Jon Snow”, “Daenerys Targaryen” and “Samwell Tarly”.

```{r}
filter(all_deaths, year_of_death <= 299)
```

```{r}
filter(all_deaths, gender == 0 & !(allegiances %in% c("Lannister", "House Lannister"))) # This is how to use the opposite of the %in% function
```

```{r}
filter(all_deaths, name %in% c("Jon Snow", "Daenerys Targaryen", "Samwell Tarly"))
```

### arrange

```{r}
arrange(all_deaths, book_intro_chapter) # arrange dataset by a certain column
```

```{r}
arrange(all_deaths, desc(book_intro_chapter))
```

```{r}
arrange(all_deaths, book_of_death, death_chapter) # can arrange by multiple columns
```

Arrange all_deaths by allegiances. What happens when you arrange by a character column?
Arrange all_deaths by allegiances and book_intro_chapter
Arrange all_deaths by descending year_of_death

```{r}
arrange(all_deaths, allegiances)
```

```{r}
arrange(all_deaths, allegiances, book_intro_chapter)
```

```{r}
arrange(all_deaths, desc(year_of_death))
```


## mutate

```{r}

mutate(all_deaths, 
       years_survived = year_of_death - 298, 
       .after = year_of_death) 

```

```{r}
mutate(all_deaths, 
       book_of_death * 5) # adds a new column, but names it "book of death * 5"
```


```{r}
mutate(all_deaths, 
       bod = book_of_death * 5) # adds a new column, but names it "book of death * 5"
```

```{r}
mutate(all_deaths, test = "test") # simply adds a new column, all of "test
```

```{r}

mutate(all_deaths, 
       year_of_death = is.na(year_of_death)) # returns true or false if they are NA for this column

```

```{r}
mutate(all_deaths, 
       nobility = as.logical(nobility))
```

```{r}
mutate(all_deaths, 
       name = as.numeric(name)) # just be wary - strings dont work well with this
```


```{r}
mutate(all_deaths, avg_death = mean(year_of_death, na.rm = TRUE)) # avergaes out all year_of_death for whole table and pops into new column
```

### summarise

```{r}
summarise(all_deaths,
          avg_death = mean(year_of_death, na.rm = TRUE),
          sum_males = sum(gender))
```

### group_by

```{r}
group_by(all_deaths, allegiances)# grouped data into 21 groupsd depending on allegiance column
```

```{r}
deaths_grouped <- group_by(all_deaths, allegiances)

```

```{r}

deaths_grouped <- group_by(all_deaths, allegiances)

summarise(deaths_grouped,
          character_count = n()) # n() for each of our groups(allegiance column), returns number of rows
```

```{r}
deaths_grouped <- group_by(all_deaths, nobility, gender)

summarise(deaths_grouped,
          char_count = n()) # return count of noble and non noble, males and females
```
### PIPE


%>% pipe operator is shift-ctrl-m

```{r}

have_died <- filter(all_deaths, !is.na(book_of_death)) # filter out NA values

died_grouped <- group_by(have_died, allegiances)
died_counts <- summarise(died_grouped, count = n())

arrange(died_counts, desc(count))

```


USING PIPE to make above code better

```{r}
all_deaths %>% 
  filter(!is.na(book_of_death)) %>% # filter out NA values in the column book_of_death
  group_by(allegiances) %>%   # group into allegaince groups
  summarise(count = n()) %>% # summarise by counting entries per group
  arrange(desc(count))      # arrange by descending count



```

### pull

```{r}
count <- all_deaths %>% 
summarise(count = n()) %>% 
pull()

count
```
```{r}
pull(count) # can pull variable values so we can use it as a single variable in other places
```

