---
title: "R Notebook"
output: html_notebook
---
```{r}
library(tidyverse)
library(janitor)
```

### Cleaning Data

First 4 lines are "meta-data" -- data about data

```{r}
video_games <- read_csv("../steam-200k-sample.csv", skip = 4) # skips the first 4 lines OF INPUT - this is if an excel file has rubbish lines at the top that we dont want to use
```



```{r}
video_games <- read_csv("../steam-200k-sample.csv", 
                        skip = 4,
                        col_types = cols("bought online" = col_logical())) # this changes this column to logical true or flase rather than 1s and zeros
video_games
```



```{r}
#read_csv("../steam-200k-sample.csv", skip = 4) %>%
# mutate(`bought online` = as.logical(`bought online`)) # need back ticks! its the key above tab for this code to work

# PROBABLY BETTER OFF USING THE ABOVE CODE CHUCK THIS SEEMS CLUNKY
```

```{r}
names(video_games)
```


```{r}

video_games_janitor_clean <- video_games %>% 
  clean_names() # this passes the data through to the clean names function which adds underscores into column names and clears out upper case etc

video_games_janitor_clean
names(video_games_janitor_clean) # this scan then be seen by calling the names function which gives us the new column names

```



```{r}
video_games_janitor_clean %>% 
  rename("game_title" = "game_name",
         "activity" = "what",
         "playing_hours" = "value")
```

```{r}

```



## Session 2 --- Missing Values

Missing values are not the same as NULL values

```{r}
library(CodeClanData)
comms_data
```

MNAR 

Fine
Analysis can accomodate this

MAR

Maybe ok
Means the missing data is related to another external variable

MNAR

Bad news
This means the data is missing due to underlying pattern within the same variable


Missing values - eg place of birth may be unknown - this is differnet from an NA

NA affects all functions tried to perform actions on it 


```{r}
NA - explicitly code missingness
NAN - eg 0/0 not a number
Inf - eg 1/0 infinte value
-Inf - eg -1/0 minus infinite
"unknown" - character representation, eg refusal to answer a question
NULL - complete lack of a value

```



```{r}
comms_data %>% 
  drop_na() # this drops all data - risks losing a lot of data!!!
```

### Imputation

this involves replacing NA values with something sensible

maybe the mean? the median? the range

```{r}

comms_data %>% 
  pull(monthly_charges) %>% # this pulls the column we want to work with
  mean(na.rm = TRUE) # this calculates the median

comms_data %>% 
  pull(monthly_charges) %>% # this pulls the column we want to work with
  median(na.rm = TRUE)

comms_data %>% 
  pull(monthly_charges) %>% 
  range(na.rm = TRUE) # this returns the min and max

min_charge <- comms_data %>% 
  pull(monthly_charges) %>% 
  min(na.rm = TRUE)

max_charge <- comms_data %>% 
  pull(monthly_charges) %>% 
  max(na.rm = TRUE)





max_charge - min_charge # this is called the range of the data

summary(comms_data) # this shows these values too

```

### Leave the NAs alone? This might be ok tho it may cause problems


```{r}
is.na(NA)

is.na(NULL)

is.na(NaN)

is.na(Inf)

is.na("unknown")
```

```{r}
comms_data
```

### standard missing values

```{r}
comms_data %>% 
  filter(is.na(monthly_charges)) # filter for the NA values
```

```{r}
comms_data %>% 
  summarize(count = sum(is.na(monthly_charges)),# counts the NA values for a column
            mean = mean(is.na(monthly_charges))) # this gives the percentage of NA values
```

## Imputation
coalesce (thing to fix, thing to fix with)



```{r}
comms_data %>%
  mutate(monthly_charges_median = coalesce(monthly_charges, median(monthly_charges, na.rm = TRUE)),
         monthly_charges_mean = coalesce(monthly_charges, mean(monthly_charges, na.rm = TRUE)),
         .after = monthly_charges) # this inserts the mean and median into new columns 
  
  
```


```{r}

```


```{r}
comms_data_imputed_values <- comms_data %>%
    mutate(monthly_charges2 = coalesce(monthly_charges, mean(monthly_charges, na.rm = TRUE)))
  
  
```


```{r}

comms_data <- clean_names(comms_data) # this cleans the names of the columns

comms_data_imputed_values <- comms_data %>% 
  mutate(monthly_charges2 = coalesce(monthly_charges,
                                     median(monthly_charges, na.rm = TRUE)), # creates a column called monthly_charge2. The NA values have been replaced with the median of the column
         monthly_charges3 = coalesce(monthly_charges,
                                     mean(monthly_charges, na.rm = TRUE)), # creates a column called monthly_charges3. The NA values from monthly_charges have been replaced with the mean of the column
         
         .after = monthly_charges) # this puts the columns after monthly charges for ease of viewing

comms_data_imputed_values
```

```{r}
ggplot(comms_data_imputed_values, aes(customer_id, monthly_charges2)) + 
  geom_point()

ggplot(comms_data_imputed_values, aes(customer_id, monthly_charges3)) + 
  geom_point()
```



### Non-Standar missing values

```{r}
comms_data
```

--

NA - proper NA

na

N/A

```{r}
comms_data %>% 
  filter(is.na(total_charges)) # finds the NAs only

comms_data <- comms_data_imputed_values %>% 
  mutate(total_charges = na_if(total_charges, "na"), # changes the column to turn "na" into a proper NA
        total_charges = na_if(total_charges, "N/A")) # changes the column to turn "N/A" into a proper NA

  
```

coalesce - turns NA into a value eg a mean

na_if turns a value into an NA eg --, N/Aa


```{r}
comms_data_cleaned <- comms_data %>% 
  mutate(payment_method = na_if(payment_method, "--"), #
         payment_method = coalesce(payment_method, "unknown")) #puts in "unknown" in payment method in place of NA values
comms_data_cleaned
```
```{r}
comms_data_cleaned <- comms_data %>% 
  mutate(total_charges = na_if(total_charges, "na"), # makes"na" a proper NA value
         total_charges = na_if(total_charges, "N/A"), # makes "N/A" a proper NA value
         payment_method = na_if(payment_method, '--'), # makes "--" a proper NA value
         payment_method = coalesce(payment_method, 'unknown')) # makes NA in payment method column into "unknown"
comms_data_cleaned
```


## Recoding

```{r}
# students_big %>% 
# (ageyears) %>% 
#  mutate(older_than_12 = ageyears > 12) # this makes a new column treu false for older than 12

students_big %>% 
select(ageyears) %>% 
  mutate(older_than_12 = if_else(ageyears > 12, "Older than 12", "12 or younger")) # this makes a new column and puts in text depedning on the if then else test

```


if_else(TEST, value if TRUE, value if FALSE)

```{r}
students_big %>% 
  select(gender) %>% 
  mutate(gender_receded = if_else(gender == "M", "Male", "Female"))

```

```{r}
students_big %>% 
  mutate(handed = if_else(handed == "R", "Right", "Left"))

```

```{r}

students_big %>% 
  select(favourite_physical_activity) %>% 
  mutate(favourite_physical_activity = if_else(favourite_physical_activity == "Football/Soccer", "Football", favourite_physical_activity))

```

### Recode()

```{r}

students_big %>% 
  select(travel_to_school) %>% 
  mutate(travel_to_school_recoded = recode(travel_to_school, 
                                           "Bus" = "bus", 
                                           "Car" = "car", 
                                           "Walk" = "walk")) # note this is old value = new value order

students_big %>% 
  count(travel_to_school)

students_big %>% 
  select(travel_to_school) %>% 
  mutate(travel_to_school_recoded = recode(travel_to_school, 
                                           "Bus" = "bus", 
                                           "Car" = "car", 
                                           "Walk" = "walk",
                                           .default = "other"))

```

```{r}

students_big %>% 
  select(ageyears) %>% 
  mutate(ageyears_recoded = recode(ageyears, "12" = 120, # changes 12 to 120, all other values to 0
                           .default = 0)) # generally dont use recode for numerical values


```



```{r}
data <- read.table(header = TRUE, text="
 subject sex control cond1 cond2 compl
       1   M     7.9  12.3  10.7  T
       2   F     6.3  10.6  11.1  T
       3   F     9.5  13.1  13.8  N
       4   M    11.5  13.4  12.9  F
")

data

```
```{r}
data %>% 
  mutate(subject = recode(subject, "1" = "001",  "2" = "002", "3" = "003", "4" = "004")) %>% 
  mutate(sex = recode(sex, "M" = "Male", "F" = "Female")) %>% 
  mutate(compl = recode(compl, "T" = "True", "F" = "False", .default = NA_character_)) # need to explicitly tell this is an NA character if other values of columns are characters


```

### Case_when

```{r}

x <- 1:10
x

case_when(x <= 5 ~ "A") # anything less than 5 gets set to 5, the rest to NAs
```
```{r}

# think of this like a sieve that we drop data through passing tests on the way
case_when(x <= 5 ~ "A", # anything less than 5 is A
          x <= 7 ~ "B", # anything less than 7 is B
          TRUE ~ "C")  # anything left over is assigned to C
x
```

```{r}
students_big %>% 
  select(ageyears) %>% 
  mutate(
    age_divisible = case_when(
      ageyears %% 2 == 0 ~ "Even",
      ageyears %% 3 == 0 ~ "Divisible by 3",
      ageyears %% 5 == 0 ~ "Divisible by 5",
      TRUE               ~ "Other number"
    )
  )




```

```{r}
students_big %>% 
  select(ageyears) %>% 
  mutate(
    ageyears = case_when(
      ageyears %% 2 == 0 ~ "Even",
      ageyears %% 3 == 0 ~ "Divisible by 3",
      ageyears %% 5 == 0 ~ "Divisible by 5",
      TRUE               ~ as.character(ageyears) # have to be explicit in the type of variable(eg character/numeric) in this case needs to be a character - this is doubly true for NA values
    )
  )
```

### Session 4 dplyr

# group by + mutate

```{r}

new_coders %>% 
  group_by(reasons_to_code) %>% 
  summarise(avg_exp_earn = mean(expected_earn)) # summarise by group by


```

```{r}

new_coders %>% 
  group_by(job_pref) %>% 
  mutate(avg_exp_earn = mean(expected_earn)) %>% 
  ungroup() # ungroup to take the groups out if going on to do more work to the dataset


```

## Window function



lag and lead functions
```{r}

chinesemeal %>% 
  mutate(last_year_cals = lag(CaloriesPerDay)) # new column with the previous years values in it


```


```{r}

chinesemeal %>% 
  mutate(last_year_cals = lead(CaloriesPerDay)) # new column with the next years values in it


```




```{r}

chinesemeal %>% 
  mutate(diff_cals = CaloriesPerDay - lag(CaloriesPerDay)) # new column with difference between the lag and the current year


```



### across function

```{r}

all_deaths

```

```{r}

all_deaths %>% 
  mutate(book1_GoT = as.logical(book1_GoT)) # changes the specified column to logical


all_deaths %>% 
  mutate(across(book1_GoT:book5_DwD, as.logical)) # applies the function "as.logical" across the given column names. column names can be entered as c("book1", "book2", ... etc). Other functions can be used instead of "as.logical"
```

