---
title: "R Notebook"
output: html_notebook
---


```{r}

is_leap_year <- function(year){
  
  if (year %% 100 == 0){
    if(year %% 400 == 0){
      return(TRUE)
    } else{
      return(FALSE)
    }
  }
  
  if(year %% 4 == 0){
    return(TRUE)
  } else {
    return(FALSE)
  }
  
  
}


```

this function tests whether a year is a leap year

the first IF checks for centenary years (ie  year divisable by 100)
IF the centenary year is also divisable by 400 then its a leap year
Otherwise the centenary year is not a lear year

the second IF checks ALL NON-centenary years (ie dont divide by 100)
IF the year is divisable by 4 its a leap year
otherwise the year is not a leap year

testthat package

```{r}
library(testthat)
```


expect_equal(): checking that the object gives the answer we expect
expect_error(): checking that the code throws an error when it should


```{r}
a <- 10
```

```{r}
expect_equal(a, 10)
```

```{r}
expect_equal(a, 11)
```
# test_that()

```{r}

test_that("Real leap year return TRUE", {
  expect_equal(is_leap_year(2020), TRUE)
  expect_equal(is_leap_year(2000), TRUE)
})





test_that("Real leap year return FALSE", {
  expect_equal(is_leap_year(3), FALSE)
  expect_equal(is_leap_year(1900), FALSE)
})


```

Task - 5 minutes
Take the following test code and write down an explanation of all the different parts:

What is the test name?

Test description is Character input returns an error, test name may be expect error

What is the expectation?

Expectation seems to be that in inputting a character there will be an error


What are you expecting this test to do?

seem to be wanting an error by inputting a character into a function that wants a numeric input.
expect error




Think carefully about the logic of this one before you write down your assumptions… you may want to try and run it to see what happens first.

# second test : one expectation

test_that("Character input returns an error", {
  expect_error(is_leap_year("year"))
})
## Test passed 🎉


test description is character input returns an erro


```{r}
test_that("Character input returns an error", {
  expect_error(is_leap_year("year"))
})

test_that("Character input returns an error", {
  expect_error(is_leap_year(2010))
})
```


```{r}

test_that("Non-numeric input returns an error", {
  expect_error(is_leap_year("year"))
  expect_error(is_leap_year(TRUE)) # R can force TURE/FALSE into year 1/0 so thats happening here and it doesnt run an error
})

```


```{r}

is_leap_year <- function(year){
  
  # add in a start condition for non-numeric data
  if(!is.numeric(year)){
    stop("Non-numeric input") # causes function to stop and kick out an error message, plus the text in the brackets
  }
  
  if (year %% 100 == 0){
    if(year %% 400 == 0){
      return(TRUE)
    } else{
      return(FALSE)
    }
  }
  
  if(year %% 4 == 0){
    return(TRUE)
  } else {
    return(FALSE)
  }
  
  
}


```


```{r}

test_that("Non-numeric input returns an error", {
  expect_error(is_leap_year("year"))
  expect_error(is_leap_year(TRUE)) # R can force TURE/FALSE into year 1/0 so thats happening here and it doesnt run an error
})

```

Write a test that checks whether 2017 and 1900 are correctly identified as not being leap years by our is_leap_year() function.

```{r}

# This test expects the function "is leap_year" to return FALSE for the 2 given inputs namely 2017 and 1900

test_that("Non leap years should return FALSE", {
  expect_equal(is_leap_year(2017), FALSE)
  expect_equal(is_leap_year(1900), FALSE)
})

```









