---
title: "R Notebook"
output: html_notebook
---

```{r}
the_name_of_my_variable <- c(3, 4, 5)

the_name_of_my_varaible == 4 

# typos cause errors, use autocorrect popup
# use smaller variable names
```

```{r}

library(fcuk)
library(tidyverse)
meen(1,2,3)
# ignore this - its suppsoed to spell check function names etc - doesnt appear to be working
```
```{r}
a = 5
```

```{r}
b = 6
```

```{r}
a = a + b
```

```{r}
a
```

Be wary that code chunks are written in correct order


```{r}

filter(the_name_of_my_variable, variable == 4) # doesnt work there is no "variable" assigned

```

```{r}

car <- mtcars

```

```{r}

library(dplyr)
library(MASS) # packages can overwrite each other functions
# order packages are loaded in changes which functions are overwritten

```
```{r}

mtcars %>% 
  dplyr::select(mpg) # if there is a conflict in function name you can specify the package like this



```
```{r}
conflicts() # lists the conflicts 
```


```{r}
mtcars %>% 
  dplyr::select(mpg, hp) %>% 
  filter(mpg < 50) %>% 
  summarise(mean = mean(hp))

# just playing about looking at error messages here  please ignore

```

Advantage of using pipes is we can run code line by line to see what works and what doesnt

```{r}


library(errorist) # automaticall copy and paste error messages into web browser
# not really much use
```
```{r}

CodeClanData::students %>% 
  group_by(favorite_school_subject, school_year) %>% 
  summarise(height_cm)

# just error checking playaround
```

```{r}
library(janitor)

mtcars$mpg == mtcars$mpg

```
```{r}


```


### MINI LAB -------------------------------------------------------------------



# Error 1.

Create some_letters which is a character vector containing the following letters: “E”, “R”, “R”, “O” and “R”.

some_letters <- ("E", "R", "R", "O", "R)


```{r}
some_letters <- c("E", "R", "R", "O", "R")
# missing a " after "R
# missing a c before the bracket
```

# Error 2.

Create message which is a vector containing the following strings: “does”, “this”, “work” and “?”, then convert the vector to uppercase using toupper.

message <- c("does", "this" "work", "?")

toupper(messege)

```{r}
message <- c("does", "this", "work", "?")

toupper(message)
# typo in message, missing comma after "this"
```
# Error 3.

Select columns mpg, cyl, and disp from mtcars and keep only rows where mpg is less than 20.

mtcars %>% 
  select(mpg, cyl, disp) %>>%
  filter(mpg < 20)
  
```{r}
mtcars %>% 
  dplyr::select(mpg, cyl, disp) %>%
  filter(mpg < 20)
# was trying to use the wrong "select" function!!! AHHHH! unload MASS package
# also double > sign in the pipe operator
```
  
# Error 4. 
 
  Select columns Sepal.Length, Sepal.Width, and Species and arrange by Sepal Length.
  
  iris %>% 
  select(Sepal.Width, Species) %>% 
  arrange(Sepal.Length)
  
```{r}
iris %>% 
  dplyr::select(Sepal.Width, Sepal.Length, Species) %>% 
  arrange(Sepal.Length)
# missing Sepal Length in the select function
```
  
# Error 5.

Find the average tooth length for where the supplement is vitamin C.
  
  ToothGrowth %>%
  summarise(avg_len_vc = mean(len)) %>% 
  filter(supp == "VC")
  
```{r}
ToothGrowth %>%
  filter(supp == "VC") %>% 
  summarise(avg_len_vc = mean(len)) 
  
# filter and summarise were the wrong way round

```
  
# Error 6. 
Create a vector of animals (cat, dog, rabbit, horse), then use the sample function to get 10 random animals from those categories.
  
  animals <- c("cat", dog, "rabbit", "horse")
  sample(animals, 10)
  
```{r}

animals <- c("cat", "dog", "rabbit", "horse")
  sample(animals, 10, replace = TRUE)

# missing "" in dog in the list
  # also the sample function needed the replace = true as the sample size 
  # wanted is bigger than the vector
```
  
# Error 7.

Create dog_names and cat_names, then use intersect to check which common pet names are shared by both dogs and cats.

dog_names <- c("Lucy", "Molly", "Charlie", "Luna", "Bella", "Cooper")

cat_names < c("Luna", "Lily", "Lucy", "Charlie", "Max", "Oliver")
Intersect(dog_names)

```{r}

dog_names <- c("Lucy", "Molly", "Charlie", "Luna", "Bella", "Cooper")

cat_names <- c("Luna", "Lily", "Lucy", "Charlie", "Max", "Oliver")

intersect(dog_names, cat_names)
# assignment operator had a typo
# intersect had a capital letter which took me ages to notice
```

# Error 8. 
Filter weights equal to or below 150 grams, then create a new column ‘weight_kg’ that contains the weight in kilograms.

chickwts %>% 
  filter(weight =< 150) %>% 
  mutate(weight_kg == weight * 0.001)
  
```{r}
chickwts %>% 
  filter(weight <= 150) %>% 
 mutate(weight_kg = weight * 0.001)
# =< instead of <=
# needed single = rather than == 
```
  
# Error 9. 
Create a tibble with column alphabet which contains the English alphabet and column numbers which has a unique number for each letter. Then remove rows where the letter is A, B or C.

tibble(alphabet = LETTERS(),
       numbers = 1:25) %>% 
  filter(letter != c(A, B, C))
  
```{r}

tibble(alphabet = LETTERS,
       numbers = 1:26) %>% 
  filter(!alphabet %in% (c("A", "B", "C"))) 
# number to 26 not 25
# no brackets after LETTERS
# ABC needed "" 
# "alphabet" not "letters" which doesn't exist as a variable

```
  
  
# Error 10.

  Clean the variable names and drop missing ratings from scent_ratings and calculate an average rating per scent.
  
  # following is data for the below
```{r}

set.seed(999)

scent_ratings <- tibble(
  iD = rep(1:10, each = 5),
  SCENT = rep(c("New book", "Cut grass", "Rose oil", "Fresh linen", "Wet dog"), 10),
  rating = sample(c(0:5, NA), 50, replace = TRUE),
  PersonAge = rep(NA, 50)
)
```
  
```{r}
scent_ratings %>% 
  janitor::clean_names() %>% 
  drop_na(1:3) %>% 
  group_by(scent) %>% 
  summarise(mean_rating = mean(rating))

# needed two colons after janitor
# cant drop all columns as one was wholly NAs so just dropped NAs from the 1st three columns
# mean at the end needed rating entered as an argument
```
  
  