---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(CodeClanData)
```

```{r}
subject_hr <- tibble( name = c("SUBJ01","SUBJ02","SUBJ03"), 
  hr_am = c(68,72,68), 
  hr_pm = c(58,89,52))

subject_hr



```

```{r}

subject_hr %>% 
  pivot_longer(cols = starts_with(("hr")),
               names_to = "measurements_of_time",
               values_to = "heartrate")


```

income <- tibble(religion = c("Agnostic","Atheist","Buddhist", "Catholic"), 
  X10k = c(12,25,3,75), 
  X10k_to_20k = c(34,14,6,60),
  X20k_to_30k = c(45,22,10,110))

income

```{r}
income <- tibble(religion = c("Agnostic","Atheist","Buddhist", "Catholic"), 
  X10k = c(12,25,3,75), 
  X10k_to_20k = c(34,14,6,60),
  X20k_to_30k = c(45,22,10,110))

income
```

```{r}
income %>%  
    pivot_longer(cols = starts_with(("X")),
               names_to = "sections_of_income",
               values_to = "count_of_section")

```



3:45
What makes the table untidy?
How could you tidy it? (You don’t need to know how to write the code, you just need to have an idea of you would tidy them up)



```{r}

messy_orders <- tibble( name = c("Joe Bloggs","Jane Smith","Penny Jones"), 
  order_info = c("XY7282/29-06-19","LO0182/15-03-19","AS6812/04-06-19"), 
  order_amount = c(58.00,125.00,75.25))

messy_orders

```
```{r}

tidy_orders <- messy_orders %>% 
  separate(order_info, 
           into = c("order_number", "date"),
           sep = "/")
tidy_orders
```

```{r}
hospital_visits <- read_csv("../data/hospitals93to98.csv")
hospital_visits
```

```{r}
hospital_visits_long <- hospital_visits %>% 
  pivot_longer(cols = starts_with("FY"),
                names_to = "year",
               values_to = "visit_number",
               names_prefix = "FY") # this strips the FY from these years values
```

```{r}

tidy_hospital <- hospital_visits_long %>% 
  pivot_wider(names_from = Field,
              values_from = visit_number) %>% 
  separate(IcdChapter, c("code", "description"), sep = "\\. ") # need the 2 \\ as . is a special character

tidy_hospital
```

```{r}
tidy_hospital %>% 
  unite(disease, c("code", "description"), sep = ":")

```

### Mini Lab

Task 1.
Load in the country_data.csv file and take a look at it. What needs changed in this to make it tidy? Which columns need made into long format, and which need spread to rows?

Task 2.
Do the appropriate pivoting of your columns using R

Task 3.
Separate the country variable so you have the name in one column and the number in another.

Task 4.
Unite the code_letter and code_numb variables so that they are both in one column, with no spaces separating them.

```{r}

country_data <- read_csv("../data/country_data.csv")

country_longer <- country_data %>% 
  pivot_longer(cols = starts_with("Y"),
               names_to = "year", 
               values_to = "number_of_people",
               names_prefix = "Y")

country_longer %>% 
  pivot_wider(names_from = "type",
               values_from = "number_of_people") %>% 
  separate(country, c("code", "country"), sep = "\\. ") %>% 
  unite(col = "code_number", c("code_letter", "code_numb"), sep = ":")
```



