---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
single_string <- "string a"
strings <- c("string a", "string b", "string c", "string d", "striiiiiiing e")

```

```{r}
pattern <- "a"
```

```{r}
single_string %>% 
  str_detect(pattern)
# looks for a given pattern ina string
```


```{r}
single_string %>% 
  str_detect(pattern, negate = TRUE)
# negate flips the answer from TURE to FLASE or vice versa
```

```{r}
strings %>% 
  str_detect(pattern)
```
```{r}
pattern <- "str"

strings %>% 
  str_detect(pattern)
```

```{r}
pattern <- "[abc]"

strings %>% 
  str_detect(pattern)

# this looks for any value containing any of the letters a,b or c
```

```{r}
pattern <- "[a-z]" # has to be a dash, not a ":"

strings %>% 
  str_detect(pattern)
# also works for a range
# this looks for any value containing any of the letters a,b or c
```

```{r}
pattern <- "[A-Z]" # is case sensitive

strings %>% 
  str_detect(pattern)
# also works for a range
# this looks for any value containing any of the letters a,b or c
```


```{r}
pattern <- "[A-Za-z]" # works for putting things side by side

strings %>% 
  str_detect(pattern)
# also works for a range
# this looks for any value containing any of the letters a,b or c
```

```{r}
pattern <- "(?i)[A-Z]" # (?i) makes it case insensitive

strings %>% 
  str_detect(pattern)
# also works for a range
# this looks for any value containing any of the letters a,b or c
```

```{r}
pattern <- "i{3}" # instances of what came bbeofre it (3 i's)

strings %>% 
  str_detect(pattern)
```

```{r}
pattern <- "i{3,}" #  this looks for 3 OR MORE i's

strings %>% 
  str_detect(pattern)
```

```{r}
pattern <- "i{3,6}" # looks for between 3 and 6 i's

strings %>% 
  str_detect(pattern)
```

```{r}
pattern <- "str|gr" # str OR gr

strings %>% 
  str_detect(pattern)
```

Task - 15 mins
Make a vector containing the following strings:

"We're having great fun learning about data analysis at CodeClan!"
"CodeClan is at 37 Castle Terrace in Edinburgh and ClockWise in Glasgow"
"Our cohort code starts with DE or DR"
"You get a great view of the castle or the national conservatoire from just outside CodeClan's door"
Write and test regular expressions which will match:

Any string mentioning CodeClan
Any string containing two consecutive capital letters
Any string not including the word "castle" or "Castle"

```{r}
strings <- c("We're having great fun learning about data analysis at CodeClan!",
                "CodeClan is at 37 Castle Terrace in Edinburgh and ClockWise in Glasgow",
                "Our cohort code starts with DE or DR",
                "You get a great view of the castle or the national conservatoire from just outside CodeClan's door")
```


```{r}
pattern <- "(?i)codeclan"
strings %>% 
str_detect(pattern)
```

```{r}
pattern <- "[A-Z]{2}"
strings %>% 
str_detect(pattern)
```



```{r}
pattern <- "castle|Castle"
strings %>% 
str_detect(pattern, negate = TRUE)
```

```{r}
pattern <- " [1-9][1-9] "
strings %>% 
str_detect(pattern)
```



? - zero or 1 # {0,1}
* - zero or more {0,}
+ - one or more {1,}

```{r}
string <- "@emailprovider.com"

pattern <- "[a-z]*@emailprovider.com"
string %>% 
  str_detect(pattern)
# this is just playing about trying things
```
```{r}
strings <- c(
  "This string has an_address@email.com in it", 
  "This one has user.name@company.co.uk", 
  "Now we've got other_person_123@server.net and my.name@runningoutofideas.com"
)

```



```{r}

# added this regex up by a process of elimination

pattern <- "[a-z0-9[:punct:]]+@[a-z[:punct:]]+"

#be careful with punctuation in a regex especially "." its ok if its in []
# otherwise escape its special meaning with "\\." 

#[:punct:] grabs all puncuation

strings %>% 
  str_extract_all(pattern)

# need to use str_extract_all as str_extract simply returns the first match in a string
# str_extract_all returns all matches in a string



```


```{r}
pattern <- "[a-z0-9[:punct:]]+@[a-z[:punct:]]+"


strings %>% 
  str_replace_all(pattern, "REDACTED")


```

Task - 15 mins

Make a vector containing the strings:

"Colin's mobile number definitely isn't 07777777777."
"A landline might be 01010 123456."
"Mhairi, Del, Aileen and Steph helped write these notes"
"I'm called your_name_here and I live in your_town_here."

Use regular expressions to:
Extract all the names
Extract the landline number
Insert your name and where you live into the fourth string

```{r}
strings <- c("Colin's mobile number definitely isn't 07777777777.",
             "A landline might be 01010 123456.",
             "Mhairi, Del, Aileen and Steph helped write these notes",
             "I'm called your_name_here and I live in your_town_here.")
```


```{r}
pattern <- "[A-Z]{1}[a-z]+" # 1 capital, followed by many lower letter

strings %>% 
  str_extract_all(pattern) %>% 
  flatten_chr()

```

```{r}

pattern <- "01[0-9]+ [0-9]*"

strings %>% 
  str_extract_all(pattern) %>% 
  flatten_chr()

```

```{r}

strings %>% 
  str_replace("your_name_here", "Neil Plenderleith") %>% 
  str_replace("your_town_here", "Glasgow")

```

```{r}
test_strings <- c(
  "abc", 
  "Abc",
  "abC",
  "s"
)


```

```{r}
pattern <- "^[a-z][a-zA-Z]*"    # ^ is starts with
pattern2 <- "[a-zA-Z]*[a-z]$"   # $ is end with
pattern3 <- "^[a-z]+$"          # starst and ends with loewr case

test_strings %>% 
  str_extract_all(pattern3)


```

```{r}

student <- c("Abigail", "Bjorn", "Caitlin")
age <- c(25, 56, 40)
emails <- c("abigal12@company.co.uk", "bsmith@gmail.co.uk", "caitlin_1980")

student_info <- tibble(student, age, emails)

student_info 

```

```{r}
student_info %>% 
  str_detect(emails, "[a-z0-9]+@") # str_detect doesnt work directly on tibbles
  
```
```{r}
student_info %>% 
  mutate(email_detect = str_detect(emails, "[a-z0-9]+@"))
  
```


```{r}

student_info %>% 
  mutate(emails = str_extract(emails, "@[a-z.]+"))
  
# this takes off the start of everyones email address eg for privacy

```

