---
title: "R Notebook"
output: html_notebook
---

# Connecting R to a database

- connect R to a postgres database omni pool

## Dont put login credentials on github or any public place

- be careful how we work with the login details

-- host name
-- username
-- database
-- port
-- password

2 main strategies for privacy

- gitignore
- store outside of a git repo altogether

Load in secret login credentials into R

```{r}
source("../secrets.R")
```

Connecting to a database

```{r}
library(RPostgres)

db_connection <- dbConnect(
  drv = Postgres(),
  host = om_host,
  port = om_port,
  dbname = om_atabase, # note this is mispelt in secrect.R so need to copy this missspelling over
  user = om_username,
  password = om_password,
  bigint = "numeric"
  
)

# now that we've created our connection object we don't need our credentials
# anymore we can remove these from our environment

rm(
  om_username,
  om_password
  
)

```

# testing our connection

```{r}
dbListTables(conn = db_connection)
```

## running queries

```{r}

employees <- dbGetQuery(conn = db_connection, statement = "SELECT * FROM employees;")
employees
```

```{r}
library(tidyverse)
```

```{r}

# can now run normal R commands on the data e.g.

employees %>% 
  filter(country == "Nigeria" & department == "Legal")

```

## Using GQL code chunks in R Notebooks

# if we 
```{sql, connection = db_connection}
SELECT * 
FROM employees
limit 10;
```

# running a query and capturing the output
```{sql, connection = db_connection, output.var = "out"}

SELECT * 
FROM employees
limit 10;
# Once we run this the output "out" will appear in the environment pane in the top right

```

Can call "out" in R:

```{r}
out %>% 
  select(first_name, last_name, department)
```


# When we are finished we should disconnect - this is good practice

```{r}
dbDisconnect(conn = db_connection)
```










