---
title: "R Notebook"
output: html_notebook
---

# Question 1.

Read in the data and investigate it, e.g. variables, variable types, dimensions etc.

```{r}
library(tidyverse)
wines <- read.csv("wine_ratings.csv")
```

```{r}
skimr::skim(wines)
glimpse(wines)
```


# Question 2.

Select columns title, taster_name, points, price, and country, then drop all NAs. Save into a variable called wine_subset. Use this variable for all subsequent tasks.


```{r}
winesna <- wines %>% 
  select(title, taster_name, points, price, country) %>% 
  drop_na()


```

# Question 3.

Find the 5 countries with the highest mean points.

```{r}
winesna %>% 
  select(country, points) %>% 
group_by(country) %>% 
summarise(country_mean = mean(points)) %>% 
slice_max(n = 5, country_mean)

winesna %>% 
  select(country, points) %>% 
group_by(country) %>% 
summarise(country_median = median(points)) %>% 
slice_max(n = 5, country_median)
```




# Question 4.

Find the 5 countries with the lowest mean prices.

```{r}

winesna %>% 
  select(country, price) %>% 
group_by(country) %>% 
summarise(price_mean = mean(price)) %>% 
slice_min(n = 5, price_mean)

winesna %>% 
  select(country, price) %>% 
  group_by(country) %>% 
  summarise(price_median = median(price)) %>% 
  slice_min(n = 5, price_median)
```




#Question 5.

Does the result differ if you use the median instead of the mean for Questions 3 and 4?

Answered above - it does seem to make a difference. the median returns either whole number or a ".5" whereas the mean returns a number with 4 decimal places so there is more scope for countries to have the same medians hence more rows are returned as ties are counted



#Question 6.

Find the names of all the tasters using distinct. Then take a random sample of 5 names using sample_n. Save these 5 names into a new variable called taster_sample which should be a character vector containing only the names.

```{r}


taster_sample <- winesna %>% 
  distinct(taster_name) %>% 
  slice_sample(n=5) %>% 
  pull()
taster_sample

```



#Question 7

Do the same thing for country, i.e. create a character vector called country_sample using distinct and pull, but this time use sample_frac to pick 10 countries at random.


```{r}
country_sample <- winesna %>% 
  distinct(country) %>% 
  slice_sample(n=10) %>% 
  pull()

country_sample
```



#Question 8

Filter the rows of wine_subset so you are left only with data from the 5 tasters in your taster_sample and from the countries in your country_sample. Save this into a new variable called wine_sample.

```{r}
wine_sample <- winesna %>% 
  filter(taster_name %in% taster_sample & country %in% country_sample)

wine_sample
```


#Question 9

Using your wine_sample, create a new column difference_from_avg which contains the difference between each wine’s rating (points) and the average rating for that taster, then arrange the differences to see the highest positive difference first. [Hint: you will first have to calculate the average points for each taster]

```{r}

wine_sample %>% 
group_by(taster_name) %>% 
mutate(taster_mean_points = mean(points), difference_from_avg = points - taster_mean_points) %>% 
arrange(desc(difference_from_avg))

  


```

