---
title: "R Notebook"
output: html_notebook
---
# if statements

```{r}

if (weather == "Sunny") { 
  hang_up_washing()
  
  
  
}

```

```{r}
x <- 10
y <- 15

if (x < y) { # then
  print("Yes")
  
}

```

```{r}
x <- 10
y <- 5:20

if (x %in% y) { # then
  print("Yes")
  
}

```

```{r}

x <- 10
y <- 5:20

if (x %in% y) { # then
  print("Yes")
  
}

```

Task - 5 minutes
Write an if statement that prints ‘Yes that divides by 2’ if a number x is divisible by 2 (hint: use %% 2 == 0 as your expression), try it out for different values of x.

```{r}
x <- 9

if (x %% 2 == 0) { #then
print("Yes that divides by 2")
  
} else {
  print("Doesnt divide by 2")
  
}

x <- 3
x <- 10



```

```{r}
x <- 6
y <- 13

if (x %in% c(3, 4, 5)){
  print("x is in our vector")
  if (x == 5){
    print("x is exactly 5")
  }
} else {
  print("Start of else statement")
  if (y %% x == 0){
    print("x divides y")
  } else{
    print("x does not divide Y")
  }
  print ("End of else statement")
}

# 
# try for 
# x-5, y-10 - Going to print x is in our list and x is exactly 5
# x=6, y=12 - Print start of else statement, print x divides y and End of else statement
# x=6, y=13 - Print start of else statement, print x does not divide y and End of else statement
# got thse wrong - remember once in an else it will do everything til that particul else closing bracket
```

## Loops

"for" loops
Used when we _know_ how many iterations we need
will end after the desired number of loops

"while" loops
Used  when we _dont_ know how many iterations
will end when a condition is met

Bear in mind loops are very useful but we can usually avoid them in R




```{r}
v

for(element in v) { # first for loop, give "element" a meaningful name to make life easier
  print(element * 2)
  
}

```

```{r}
weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")

for(day in weekdays){
  phrase <- paste("The day is", day)
  print(phrase)
}

```

# While Loop

Sometimes you can get stuck in an infinite loop

```{r}
current_value <- 1 # initialise your variables first
stop_value <- 10

while(current_value < stop_value){
  current_value <- current_value + 1
  print(current_value)
}

```


### Alternative ways to index for a loop


```{r}

len_weekdays <- length(weekdays)

for(i in 1:len_weekdays){
  phrase <- paste("the day is", weekdays[i])
  print(phrase)
  
}
```

### Build a vector from a loop __TERRIBLE PRACTICE__

```{r}

weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
weekdays
len_weekdays <- length(weekdays)

for (i in 1:len_weekdays) {
  plural <- paste0(weekdays[i], "s")  # creating a new value from the old value on the index i
  weekdays[i] <- plural               # modify each element based on the index i
  
}
weekdays
```

### Nested loops

```{r}

m <- matrix(1:12, 4)
nrows <- nrow(m)
ncols <- ncol(m)

# for (row_i in 1:nrows){
#   print(m[row_i, ])
#   
# }

for (row_i in 1:nrows) {
 for(col_i in 1:ncols) {
   print(m[row_i, col_i])
 }
}
```




```{r}
product_matrix <- matrix(NA, nrow = 5, ncol = 5)
```

```{r}
# this fills the matrix with the product of the row and column numbers

product_matrix <- matrix(NA, nrow = 5, ncol = 5)

nrows <- nrow(product_matrix)
ncols <- ncol(product_matrix)

# for (row_i in 1:nrows){
#   print(m[row_i, ])
#   
# }

for (row_i in 1:nrows) {
 for (col_i in 1:ncols) {
   product_matrix[row_i, col_i] <-  row_i*col_i
 }
}
product_matrix
```

```{r}
example_matrix = matrix(1:25, nrow = 5, ncol = 5)
example_matrix 

for (row_index in 1:5){
  for (col_index in 1:5){
    if (row_index == col_index){ # logical test within the 2 for loops
      print(example_matrix[row_index, col_index])
    }
   }
}
```



```{r}



```
















