---
title: "R Notebook"
output: html_notebook
---

## Lists

```{r}
library(tidyverse)
```

```{r}
x <- list(1, "a", TRUE, c(1, 2, 3)) # list allows various datatypes

x

vector <- 1:5
vector

as.list(vector) # can change a vector to a list
```
LIST OUTPUT BELOW
[[1]] element 1
[1] 1   has 1 element - a 1

[[2]]
[1] "a"

[[3]]
[1] TRUE

[[4]]
[1] 1 2 3

## Subset

```{r}
x[[2]] returns the vector in the second position in a list
x[2]   returns the second element AS A LIST
```

```{r}
glimpse(x)
```
## Named list

```{r}

named_list <- list(a = "I love R", b = 1:10) # this names the elements of a list
named_list

glimpse(named_list)

named_list$a # extracts the named element "a"
```

```{r}

names(mtcars)

mtcars[1]
mtcars[[1]]

mtcars %>% 
  select(mpg) # same as single bracket above

```

### Nested List

```{r}

student_info <- list(
                "Agnes" = list(height = 1.8, shoe_size = 8L),
                "Bobby" = list(height = 2.1, gender = "Male"),
                "Charlie" = list(gender = "Female", likes_beets = TRUE)
)

student_info

str(student_info) # gives info about the list

# take out agnes height using the "$"
student_info$Agnes$height # use if it wont change



student_info[["Agnes"]][["height"]] # use this if you want ot access later or with variables, is also safer if names etc change

n <- "Agnes"
student_info[[n]][["height"]] # works with variable names

```
# Task

    Make a named list that contains a numeric vector and a character vector.
    Access the second element of your list using double square brackets [[]]. What is it’s class?
    Access the second element of your list using single square brackets []. What is it’s class?
    Access the second element of your list by name. What is it’s class?


```{r}

neil_list <- list(a = 1, b ="Jim")

neil_list[[2]] %>% 
class() # returns character

neil_list[2] %>% 
class() # type of list 

neil_list$b %>% 
class() # returns vcharacter SO SAME AS DOUBLE BRACKETS

neil_list[["b"]] %>% 
class() # returns character SO SAME AS DOUBLE BRACKETS


```

## Matrices

all the same character types

```{r}

matrix(1:4) # puts it into 1 column - COLUMN ORDER

```

```{r}

matrix(1:4, nrow = 2) # puts it into 2 rows

```



```{r}

matrix(1:12, ncol = 3) # puts it into 3 column

```


```{r}

m <- matrix(1:12, ncol = 3, byrow = TRUE) # pforced into putting ito wors first
m

class(m) # type matrix
typeof(m)

```



### Index a matrix

```{r}
m <- matrix(1:12, ncol = 3)

m[1:2, 1:3] # select columns 1-3, rows 1-2


```

```{r}

m[1, ] # gives the first row (NOTE THIS IS A VECTOR)

m[ , 1] # gives the first column (AGAIN AS A VECTOR)



```
```{r}

m[1, 2] # export the 1st row, second column value

```

```{r}

m[2:3, ] # if asking for multiple rows and columns then will export as a matrix still

```

```{r}
m[c(1,3), c(1,3)] # for non consecutive rows add them into a c()
```
```{r}
m[ ,-2] # drops the second column
```

```{r}
v <- 10:15
v
v * 2
```
Matrices work in a vectorised way
```{r}
m * 2 # mulitply all by 2
m - 2 etc
```

```{r}
rowSums(m) # sum the rows

colSums(m) # sum the columns

rowMeans(m) # mean of the rows

colMeans(m) # mean the columns
```

Task - 5 minutes
    Make a matrix with 3 rows and 4 columns. Each column should contain the numbers 1 to 3.
    Find the element in the 2nd row, 3rd column.
    Check that all the columns are identical by checking that their sums are the same.
    
```{r}

npmatrix <- matrix(1:3, nrow = 3, ncol = 4)
npmatrix

#useful:
rep(1:3, 3)
rep(1:3, each = 3)
```

```{r}

npmatrix[2, 3]

```
```{r}

colSums(npmatrix)

```
```{r}




colnames(m) <- paste0("col_", 1:3) #adds colnames, paste0 doesnt add a separator between things to join
m
rownames(m) <- paste0("row_", 1:4)

m[, c("col_1", "col_2")]
```

