---
title: "R Notebook"
output: html_notebook
---

# Joins

4 mutating join verbs:
* inner_join() : returns all rows from two tables with matching records.
* left_join() : returns all records in left table, and any matching records in another.
* right_join() : returns all records in right table, and any matching records in another.
* full_join() : returns all records in both tables, matching where possible.

2 filtering join verbs:
* semi_join() : returns only rows from the first table where one or more matches are found in the second table. Unique only
* anti_join() : returns only rows from the first table where no matches are found in the second table, unique only

2 binding verbs:
* bind_rows() : add rows to a table
* bind_cols() : add columns to a table



```{r}
library(tidyverse)
```


```{r}

customers <- tibble(
  customer_ID = c(001, 002, 004, 005, 008, 010), 
  name = c("John Smith", "Jane Adams", "Robert Landry", "Jane Tow", "Raul Tann", "Hilary Joyal"),
  email_address = c("johnsmith@gmail.com", "janea@gmail.com", "rlabdry@hotmail.com", "janet89@aol.com", "paul.tann1@gmail.com", NA),
  shipping_address = c("32 Station Road, Edinburgh", "42 Park Drive, Edinburgh", NA, "10 Gardiner Rd, Edinburgh", "12 Main St, Edinburgh", " 234 Queensferry Rd, Edinburgh,")
)

customers
```

```{r}
orders <- tibble(
  customer_ID = c(001, 002, 005, 010, 012, 001), 
  order_number = c("KU65781", "SU98201", "JS76821", "YR28212", "JS91221","YR21873"),
  order_total = c("£32.50", "£56.00", "£100.02", "£20.25", "£39.50", "£45.00"),
  product_code = c("XZY780", "XYZ123", "XYZ256", "XYZ262", "XYZ345", "XYZ127")
)

orders 
```

# Trying joins

```{r}

# general syntax:
# join(datax, datay, c(keyx, keyy))

```

```{r}

inner_join(customers, orders, "customer_ID")
# returns all rows from two tables with matching records.
```

```{r}

# possibly the more popular function to use

left_join(customers, orders, "customer_ID")

#includes all values from the left table, matching recrods from the right table
```

```{r}

right_join(customers, orders, "customer_ID")

# all values from the right table, only matching records from the left
```

```{r} 

full_join(customers, orders, "customer_ID")

```

# Join with multiple values
```{r}
# arbitrarily rename our customer_ID column for the customer data
customers_adj <- rename(customers, ID = customer_ID)

# now let's join that 
inner_join(customers_adj , orders, by = c("ID" = "customer_ID"))
# order of arguments after "by =" matters here - same order as tables
```


```{r}

customers2 <- customers %>% 
  mutate(payment_type = c("card", "cash", "cash", "card", "card", "card"))

orders2 <- orders %>% 
  mutate(payment_type = c("card", "cash", "cash", "card", "card", "card"))


```


```{r}
inner_join(customers2, orders2, by = c("customer_ID", "payment_type"))
```

# Filtering Joins

```{r}

semi_join(customers, orders, "customer_ID")

# returns only rows from the first table where one or more matches are found in the second table. Unique only
```

```{r}

anti_join(customers, orders, "customer_ID")

# returns only rows from the first table where no matches are found in the second table, unique only
```


# Binding Joins


```{r}
statuses <- tibble(
  status = c("active", "active", "inactive", "inactive", "active", "inactive")
) 
```

```{r}

bind_cols(customers, statuses)
# bind a column onto the end of dataframe
# not used too much
```

```{r}

past_customers <- tibble(
  customer_ID = c(020, 021), 
  name = c("Dave Brown", "Jess Bishop"),
  email_address = c("dbrown@hotmail.com", "jessb@gmail.com"),
  shipping_address = c("57 Park View, Leith", "12 Reid Way, Broxburn")
)

past_customers

```

```{r}

bind_rows(customers, past_customers)

# adds rows onto the bottom of a tibble/dataframe
# column names need to be the same or it will make new columns

```

