---
title: "R Notebook"
output: html_notebook
---

ctrl alt I - insert code chunk

```{r}

2+5

4+9

```
```{r}

```

Ctrl-shift-enter - run whole chunk
ctrl-enter-run line cursor is on / run selected code

ctrl - shift - P command pallet - lists commands and shortcuts

```{r}

mtcars

```


### Data Types




Categories
Nominal
  discrete - Eg Texas
  ordinal - 1st, 2nd , 3rd etc. Season of the year, things rated from 1st to 10th. An implicit order
  
  
Numeric
  Ratio
  Interval
  Continous - eg height 3.445556677777 etc
  Discrete - no inbetween - whole integers how many people live in your home





```{r}

dbl1 <- c(1,4,6,3)
dbl2 <- c(1, 4, 6, 3, "x")
ingr1 <- c(1L, 4L, 6L, 9L)

class(dbl1) # continous varaibles
class(dbl2) # Discrete values
class(ingr1)

typeof(dbl1)
typeof(dbl2)
typeof(ingr1)



char1 <- c("a", "b", "a", "c")
char2 <- factor(char1)


char1
char2

typeof(char1)
typeof(char2)

class(char2)

# logical

typeof(c(TRUE, FALSE, TRUE, FALSE))


```

R deals with different types of data in a way that is appropriate for each type


```{r}

"jimmy"
""

```

